|

### 2020-05-19 paper/Manuscript.Rmd
Rewrite intro

The (social) sciences are amidst a paradigm shift towards open science. 
Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

This paper introduces WORCS: A Workflow for Open Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible. WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals. The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an RStudio project template.

@@

The remainder of the paper is structured as follows. ... <!-- AL: Or some variation of this obligatory end-of-intro paragraph. As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here. -->

# Defining open science

Although Open Science is advocated by many, there is no clear definition of Open Science (LOEK? ref multiple definitions).
The TOP-guidelines have been influential as a concrete operationalisation of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement [REF Loek paper].
This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions:
If such requirements exist, this workflow will help fulfill them.

@@

Sharing all research output also provides a proper basis for research evaluation according to the San Francisco Declaration on Research Assessment (DORA; https://sfdora.org/), which increasingly plays a role in grant funding and hiring or promotion procedures. Direct access to research objects allows funding bodies and hiring committies to evaluate research quality based on content rather than relying on spurious surrogate indicators like journal impact factors, conference rankings, and h-indexes. The detailed version control and commit tracking of github and similar platforms furthermore make it possible to assess the relative contributions made by different researchers.

The FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] advocate that research data should be 
Findable, Accessible, Interoperable and Reusable. Moreover, these principles are
increasingly applied to other types of research output 
[@lamprechtFAIRPrinciplesResearch2019]. Research
organizations and funders use the FAIR principles as standard for open, 
transparent and reproducible science.
The WORCS workflow helps users meet the FAIR principles by making all research objects -
including (origina) data and code - accessible in a (public) repository. 
The `worcs` package additionally
facilitates meeting the FAIR principles in several ways. First, each `worcs` project has a YAML file
(https://yaml.org/) that lists the research objects in the project, which will allow web crawlers 
to index `worcs` projects. Second, a codebook is automatically generated for each data file stored
using `open_data()` or `closed_data()`. We encourage users to elaborate on the default codebook by adding
variable description and categories, which might - in the future - enable indexing data by topic area. 
Relatedly, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.

>>>>

The (social) sciences are arguably past the tipping point of the paradigm shift
towards open science.
Several immediate causes stirred up support for this transition; including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science is not merely a cure for this crisis - it is also an
opportunity [@adolphOpenBehavioralScience2012].
Open science helps researchers to more easily conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012].
Capitalizing on these advances has the potential to accellerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many individual researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
adopt open science practices can be daunting. Making the transition effectively
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by providing a simple workflow 
that meets most requirements open science, based on best practices, and 
supported by user-friendly software in R: The Worfkflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility. The workflow (Figure \@ref(fig:workflow)) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project Open and Reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used to produce work in accordance with best 
practices in the absence of top-down support. The WORCS workflow is supported by
an R-package with user-friendly support functions, and an RStudio project
template that is, effectively, a "one click solution". This paper introduces
the WORCS workflow, discusses how it meets best practices for open science, and
illustrates the use of the `worcs` package.

@@

<!-- AL: The remainder of the paper is structured as follows. ... Or some variation of this obligatory end-of-intro paragraph. As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here. 
CJ: I was always taught that you should not explain your structure in the paper - you should just structure the paper. If the structure is unclear, make it better. -->
<!--openness, transparency, and reproducible science -->

# Meeting the requirements of Open Science

Although Open Science is advocated by many, it does not have a unitary definition.
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential concrete operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: Citation (1),
sharing (2-5), and preregistration (6-7).

WORCS meets the TOP-guidelines in a way that is amenable to the 
FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
data should be Findable, Accessible, Interoperable and Reusable. However, they are
increasingly applied as a standard for 
to other types of research output as well
[@lamprechtFAIRPrinciplesResearch2019].
WORCS helps users meet the FAIR principles by making all research objects -
including (origina) data and code - accessible in a (public) repository. 
The `worcs` package additionally
facilitates meeting the FAIR principles in several ways. First, each `worcs`
project has a YAML file (https://yaml.org/) that lists the research objects in
the project, which will allow web crawlers 
to index `worcs` projects. Second, a codebook is automatically generated for
each data file stored
using `open_data()` or `closed_data()`. We encourage users to elaborate on the
default codebook by adding
variable description and categories, which might - in the future - enable
indexing data by topic area. 
Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.

Sharing all research output also provides a proper basis for research evaluation
according to the San Francisco Declaration on Research Assessment (DORA;
https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
and promotion procedures. Direct access to research objects allows stakeholders
to evaluate research quality based on content rather than relying on spurious
surrogate indicators like journal impact factors, conference rankings, and
h-indexes. The detailed version control and commit tracking of 'GitHub' and
similar platforms furthermore make it possible to assess the relative
contributions made by different researchers.

@@

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, <!-- edit -->
The formation of "Open Science Communities" at universities around the globe
illustrates the momentum of this grassroots movement
[@brinkmanWidescaleAdoptionOpen2020].

@@
